\textbf{Supplementary materials: Development of outdoor air pollution models for estimating exposure in the Barcelona Life Study Cohort (BiSC)}

### 2023-12-20 manuscript_BiSC_outdoor_AP.tex

Air pollution is the leading environmental risk factor for human health worldwide linked to wide range of adverse health outcomes. A detailed and comprehensive exposure assessment of outdoor air pollution considering both spatial and temporal variability is fundamental to conduct epidemiological studies.

@@

Air pollution stands as the primary environmental risk factor for human health and the fourth most deadly health risk worldwide, playing a significant role in the global burden of disease \cite{cohen2017, he2020state}. Exposure to ambient air pollution has been linked with a extensive array of adverse health outcomes \cite{boogaard2022, guxens2022hei, haddad2023}. In this context urbanization have been playing a major role increasing air pollution exposure as cities grow and more population moves to urban areas \cite{nieuwenhuijsen2016}. This trends present challenges for epidemiological studies \cite{tonne2017}, where accurate estimates capturing both the spatial and temporal variability are essential to evaluate the health impacts of air pollution in urban areas \cite{boogaard2022}. \\

Over the past few decades, several modelling techniques have been adopted in environmental epidemiology to determine exposure concentrations of air pollution \cite{hoek2017, di2019no2, di2019pm25, stafoggia2019, stafoggia2020}. Among these, Land use regression (LUR) and dispersion models (DM) are the most common approaches for estimating air pollution exposure concentrations in epidemiological studies \cite{gulliver2015, dehoogh2014}. LUR models employ regression techniques to predict air pollution levels at non-monitored locations, drawing on observed air pollution data and spatial explanatory variables that work as proxies for traffic-related air pollution \cite{briggs1997, hoek2008}. In contrast, DMs rely on mathematical simulations to estimate air pollution levels. These predictions are based on data from emissions sources, chemical transformation models, pollutant dispersion models, topography, roughness, and meteorological data \cite{hoek2017}. Although these methods have been extensively used they come with inherent limitations related to lack of temporal and spatial variation. To overcome the shortcomings, hybrid models (HM) have emerged, blending both spatial and temporal advantages, leading to refined and enhanced estimations \cite{hoek2017, korek2017, tularam2021, oh2021}. \\

While significant advancements have been made in spatiotemporal modelling of air pollution over large-scale domains, such models (multi-country and national domain models) often compromise their accuracy at smaller scales \cite{dehoogh2016, chen2019, chen2020, shen2022}. This gap underscores the need for a comprehensive assessment of outdoor air pollution exposure in smaller domains to understand the health impacts of air pollution exposure, especially during critical stages of development, due to the potential lasting effects of air pollution over the lifespan \cite{selevan2000, wick2010, ghosh2021}. \\

Urban areas add a layer of complexity to air pollution modelling due to complex spatiotemporal patterns that difficult accurate estimation of ground-level air pollutant concentrations \cite{sokhi2022}. In response to these challenges, machine learning (ML) algorithms have been increasingly used, particularly decision trees-based models \cite{liu2022}. This family of models is often used due to its ability to model potential, nonlinear interaction between predictor variables \cite{liu2022treebased}. Among the different ML algorithms, random forest (RF) has emerged as one of the most robust techniques for predicting outdoor air pollution concentrations \cite{chen2019, chen2020, stafoggia2019, stafoggia2020, schneider2020, mila2023}. Furthermore, the COVID-19 pandemic has introduced unprecedented challenges for air pollution modelling and exposure assessment. These challenges encompass changes in air pollution patterns, constraints on data collection due to health and safety restrictions (lockdowns), and the complexity of data integration due to changes in traffic, mobility, and land use dynamics in cities \cite{gonzalez2022, querol2021}. \\

@@

% Put the figure text closer to the Figure 1
\captionsetup[figure]{skip=-4pt}
% We add the figure with the study domain 
\begin{figure}[!htb]
\includegraphics[width=1.0\textwidth]{figures/fig1_op2.png}
\caption{Study domain and spatial distribution (BiSC participants, BiSCAPE campaign, monitoring stations, and traffic points)}
\label{fig1}
\end{figure}

@@

%%% ---------------------------------------------------------------- %%%
%%% Table 1. Outdoor air pollution concentration BiSC-home, BiSCAPE  %%%
%%% ---------------------------------------------------------------  %%%
\begin{table}[h!]
\centering
\caption{Outdoor air pollution concentration distribution for the BiSC-home and BiSCAPE campaigns (2018-2022)}
\label{table1} 
\begin{tabular}{lllllllllll} % each one indicates a new column in the table 
\toprule
Campaign & Pollutant & Year  & mean & sd & min & p25 & p50 & p75 & max \\ 
\midrule
BiSC-home, BiSCAPE & NO\textsubscript{2} (\mu\)g/m\textsuperscript{3}) & 2018-2021 & 38.8 & 12.4 & 6.85 &30.0 & 37.5 & 45.8 & 97.3 \\ 
\multirow{5}{*}{BiSCAPE} 
& PM\textsubscript{25} (\(\mu\)g/m\textsuperscript{3}) & 2021-2022  & 13.4 & 6.14 & 5.01 & 9.06 & 11.3 & 16.2 & 35.8 \\ 
& BC (\(\mu\)g/m\textsuperscript{3}) & 2021-2022  & 1.27 & 0.50 & 0.40 & 0.88 & 1.22 & 1.52 & 2 .95 \\ 
& Fe (\(\mu\)g/m\textsuperscript{3}) & 2021-2022  & 0.25 & 0.13 & 0.07 & 0.15 & 0.21 & 0.32 & 0.77 \\ 
& Cu (ng/m\textsuperscript{3}) & 2021-2022 & 7.10 & 3.82 & 1.78 & 4.37 & 6.36 & 8.98 & 22.7 \\ 
& Zn (ng/m\textsuperscript{3}) & 2021-2022  & 40.7 & 36.5 & 6.71 &  18.9 & 28.2 & 46.7 & 273 \\ 
\bottomrule
\end{tabular}
\end{table}

The spatial distribution of the outdoor air pollution sampling campaigns is shown in \Cref{fig1}. Figure 1A illustrates a high density coverage in the central area of Barcelona, where the BiSC-home campaign was conducted. Figure 1B exhibit a similar pattern of spatial distribution but with a reduced density in sampling locations.

@@

Since different approaches were used to built outdoor air pollution models were not directly comparable due to differences on temporal basis.

@@

\textbf{\cref{fig2}} presents the distribution of the observed and estimated NO\textsubscript{2}, PM\textsubscript{2.5} and BC concentrations from the LUR, DM, and HM. The data shown are long-term estimated concentrations of pollutants, after averaging weekly estimates for all the study period (2019 - 2021).

@@

% Put the figure text closer to the Figure 3
\captionsetup[figure]{skip=6pt}
% We add the figure with the study domain 
\begin{figure}[!htb]
\includegraphics[width=1.0\textwidth]{figures/HEX_all.png}
\caption{Comparison/agreement between outdoor air pollution model estimates \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}.}
\end{figure}

>>>>

Air pollution is the leading environmental risk factor for human health worldwide linked to wide range of adverse health outcomes. A detailed and comprehensive exposure assessment of outdoor air pollution considering both spatial and temporal variability in urban areas is essential to identify critical windows of susceptibility during pregnancy to traffic-related air pollution. Within the framework of the Barcelona Life Study Cohort (BiSC), we developed Land use regression (LUR), dispersion (DM), and hybrid (HM) models to estimate outdoor air pollution exposure at residential addresses during pregnancy for several traffic-related air pollutants; NO\textsubscript{2}, PM\textsubscript{2.5}, BC, and PM\textsubscript{2.5} constituents (Fe, Cu, Zn).  The overall aim is to obtain accurate estimates of traffic-related air pollution, intended to be used in the prenatal exposure assessment for a cohort of pregnant women in the city of Barcelona.
We create 101 spatial predictor variables

@@

Air pollution stands as the primary environmental risk factor for human health and the fourth most deadly health risk worldwide, playing a significant role in the global burden of disease \cite{cohen2017, he2020state}. Exposure to ambient air pollution has been linked with many adverse health outcomes \cite{boogaard2022, guxens2022hei, haddad2023}. In this context, urbanization has been playing a major role in increasing air pollution exposure as cities grow and more population moves to urban areas \cite{nieuwenhuijsen2016}. These trends present challenges for epidemiological studies \cite{tonne2017}, where accurate estimates capturing both the spatial and temporal variability are essential to evaluate the health impacts of air pollution in urban areas \cite{boogaard2022}. \\

Over the past few decades, several modeling techniques have been adopted in environmental epidemiology to determine exposure concentrations of air pollution \cite{hoek2017, di2019no2, di2019pm25, stafoggia2019, stafoggia2020}. Among these, Land use regression (LUR) and dispersion models (DM) are the most common approaches for estimating air pollution exposure concentrations in epidemiological studies \cite{gulliver2015, dehoogh2014}. LUR models employ regression techniques to predict air pollution levels at non-monitored locations, drawing on observed air pollution data and spatial explanatory variables that work as proxies for traffic-related air pollution \cite{briggs1997, hoek2008}. In contrast, DMs rely on mathematical simulations to estimate air pollution levels. These predictions are based on data from emissions sources, chemical transformation models, pollutant dispersion models, topography, roughness, and meteorological data \cite{hoek2017}. Although these methods have been extensively used they come with inherent limitations related to lack of temporal and spatial variation. To overcome the shortcomings, hybrid models (HM) have emerged, blending both spatial and temporal advantages, leading to refined and enhanced estimations \cite{hoek2017, korek2017, tularam2021, oh2021}. \\

While significant advancements have been made in spatiotemporal modelling of air pollution over large-scale domains, such models (multi-country and national domain models) often compromise their accuracy at smaller scales \cite{dehoogh2016, chen2019, chen2020, shen2022}. This gap underscores the need for a comprehensive assessment of outdoor air pollution exposure in smaller domains to understand the health impacts of air pollution exposure, especially during critical stages of development, due to the potentially lasting effects of air pollution over the lifespan \cite{selevan2000, wick2010, ghosh2021}. \\

Urban areas add a layer of complexity to air pollution modelling due to complex spatiotemporal patterns that make difficult accurate estimation of ground-level air pollutant concentrations \cite{sokhi2022}. In response to these challenges, machine learning (ML) algorithms have been increasingly used, particularly decision trees-based models \cite{liu2022}. This family of models is often used due to its ability to model potential, nonlinear interaction between predictor variables \cite{liu2022treebased}. Among the different ML algorithms, random forest (RF) has emerged as one of the most robust techniques for predicting outdoor air pollution concentrations \cite{chen2019, chen2020, stafoggia2019, stafoggia2020, schneider2020, mila2023}. Furthermore, the COVID-19 pandemic has introduced unprecedented challenges for air pollution modelling and exposure assessment. These challenges encompass changes in air pollution patterns, constraints on data collection due to health and safety restrictions (lockdowns), and the complexity of data integration due to changes in traffic, mobility, and land use dynamics in cities \cite{gonzalez2022, querol2021}. \\

@@

% Put the figure text closer to the Figure 1
\captionsetup[figure]{skip=0pt}
% We add the figure with the study domain 
\begin{figure}[!htb]
\includegraphics[width=1.0\textwidth]{figures/figure_campaigns.png}
\caption{Study domain and spatial distribution of the monitoring campaigns}
\label{fig1}
\end{figure}

@@

%%% ---------------------------------------------------------------- %%%
%%% Table 1. Outdoor air pollution concentration BiSC-home, BiSCAPE  %%%
%%% ---------------------------------------------------------------  %%%
\begin{table}[h!]
\centering
\caption{Outdoor air pollution concentration distribution for the BiSC-home and BiSCAPE campaigns (2018-2022)}
\label{table1} 
\begin{tabular}{lllllllllll} % each one indicates a new column in the table 
\toprule
Campaign & Pollutant & Year  & mean & sd & min & p25 & p50 & p75 & max \\ 
\midrule
BiSC-home, BiSCAPE & NO\textsubscript{2} (\mu\)g/m\textsuperscript{3}) & 2018-2021 & 38.8 & 12.4 & 6.85 &30.0 & 37.5 & 45.8 & 97.3 \\
\hline
\multirow{5}{*}{BiSCAPE} 
& PM\textsubscript{2.5} (\(\mu\)g/m\textsuperscript{3}) & 2021-2022  & 13.4 & 6.14 & 5.01 & 9.06 & 11.3 & 16.2 & 35.8 \\ 
& BC (\(\mu\)g/m\textsuperscript{3}) & 2021-2022  & 1.27 & 0.50 & 0.40 & 0.88 & 1.22 & 1.52 & 2 .95 \\ 
& PM\textsubscript{2.5 - Fe} (\(\mu\)g/m\textsuperscript{3}) & 2021-2022  & 0.25 & 0.13 & 0.07 & 0.15 & 0.21 & 0.32 & 0.77 \\ 
& PM\textsubscript{2.5 - Cu} (ng/m\textsuperscript{3}) & 2021-2022 & 7.10 & 3.82 & 1.78 & 4.37 & 6.36 & 8.98 & 22.7 \\ 
& PM\textsubscript{2.5 - Zn} (ng/m\textsuperscript{3}) & 2021-2022  & 40.7 & 36.5 & 6.71 &  18.9 & 28.2 & 46.7 & 273 \\ 
\bottomrule
\end{tabular}
\end{table}

The spatial distribution of the outdoor air pollution sampling campaigns is shown in \Cref{fig1}. Figure 1A illustrates a high density coverage in the central area of Barcelona, where the BiSC-home campaign was conducted. Figure 1B exhibit a similar pattern of spatial distribution of the monitoring campaign but with a reduced density in sampling locations.

@@

Since different approaches and sampling campaigns were used to built outdoor air pollution models were not directly comparable.

@@

\textbf{\Cref{fig2}} presents the distribution of the observed and estimated NO\textsubscript{2}, PM\textsubscript{2.5} and BC concentrations from the LUR, DM, and HM. The data shown are long-term estimated concentrations of pollutants, after averaging weekly estimates for all the study period (2019 - 2021).

@@

% Put the figure text closer to the Figure 3
\captionsetup[figure]{skip=6pt}
% We add the figure with the study domain 
\begin{figure}[!htb]
\includegraphics[width=1.0\textwidth]{figures/final_HEXBIN.png}
\caption{Comparison between outdoor air pollution model estimates \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC} and \textit{PM$_{2.5}$ constituents (Fe, Cu, Zn).}}
\end{figure}

### 2023-12-20 supplementary_BiSC_outdoor_AP.tex

% Useful packages
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{authblk} % this is for multiple autho  rs 
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{cleveref}
\usepackage{booktabs}
\usepackage{multirow}

% Configure for "Fig.":
\crefname{figure}{Fig.}{Figs.}
\Crefname{figure}{Fig.}{Figs.}

@@

\subsection{Performance metrics}
The following metrics were used to evaluate the LUR model for all the air pollution assessed.

@@

% --- mean bias (MB) --- % 
\begin{equation}
\text{Mean Bias} = \frac{1}{n} \sum_{i=1}^{n} (y_i - \hat{y}_i)
\end{equation}

@@

@@

@@

%%% --- Table with performance metrics for Hybrid models --- %%%
\begin{table}[ht]
\centering
\caption{Performance metric for hybrid models.}
\begin{tabular}{llllll}
\toprule
\textbf{Dataset campaign} & \textbf{Pollutant} & \textbf{N} & \textbf{R\textsuperscript{2} (10-CV)} & \textbf{RMSE (10-CV)} \\
\midrule
BiSC-home & NO\textsubscript{2} & 1232 & 0.64 & 7.48 \\
\multirow{5}{*}{BiSCAPE} & PM\textsubscript{2.5} & 161 & 0.66 & 3.45 \\
& BC & 74 & 0.86 &  0.23 \\
& Fe & ... & 0.54 & ... \\
& Cu & ... & 0.70 & ... \\
& Zn & ... & 0.42 & ... \\
\bottomrule
\end{tabular}
\label{tablesX}
\end{table}

\subsection{Variable importance}

%%% --- Figure. Comparison between measurements and predictions values in the test dataset --- %%%
% Put the figure text closer to the Figure 3
\captionsetup[figure]{skip=6pt}
% We add the figure with the study domain 
\begin{figure}[!htb]
\includegraphics[width=1.0\textwidth]{figures/fig_HM_allmodels_importance.png}
\caption{Comparison between measurements and prediction values for the hybrid models for \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}, \textit{Fe}, \textit{Cu} and \textit{Zn}.}
\label{fig2}
\end{figure}

@@

%%% --- Figure. Comparison between measurements and predictions values in the test dataset --- %%%
% Put the figure text closer to the Figure 3
\captionsetup[figure]{skip=6pt}
% We add the figure with the study domain 
\begin{figure}[!htb]
\includegraphics[width=1.0\textwidth]{figures/fig_HM_test_all_models.png}
\caption{Comparison between measurements and prediction values for the hybrid models for \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}, \textit{Fe}, \textit{Cu} and \textit{Zn}.}
\label{fig2}
\end{figure}

>>>>

% Useful packages
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{authblk} % this is for multiple autho  rs 
\usepackage[colorlinks=true, allcolors=black]{hyperref}
\usepackage{cleveref}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{threeparttable} % For footnotes in tables
\usepackage{graphicx}  % For including graphics
\usepackage{caption}   % For customizing captions
\usepackage{cleveref}  % For enhanced referencing

% Configure for "Fig.":
\crefname{figure}{figure}{figures}
\Crefname{figure}{Figure}{Figures}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Predictor variables and sources of information --- %%% 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Predictor variables and sources of information}
To develop LUR models, we calculate a set of spatial variables that could hep us to explain pollutant concentration distribution across the study area. We calculate a total of 101 predictor variables, see \textbf{\Cref{Table S1}}. To calculate the predictor variables we applied a wide range of local and regional GIS datasets, see \textbf{\Cref{Table S2}}.

%%%% ----------------------------------------------------------------------------------%%%%
%%%% --- Table S1. Predictor variables categories created for LUR model development --- %%%
%%%% ----------------------------------------------------------------------------------%%%%

\renewcommand{\thetable}{S\arabic{table}}
\begin{table}[ht]
\centering
\caption{Predictor variables categories created for LUR model development.}
\begin{tabular}{lll}
\toprule
\textbf{Traffic} & \textbf{Land-uses} & \textbf{Urban configuration} \\
\midrule
Road length & Residential & Population \\
Major road length & Industry & Buildings density \\
Distance to road & Natural & Imperviousness \\
Distance to major road & Urban green & Slope \\
Traffic intensity & Road density & Buildings height \\
Traffic intensity only major roads &  & Street bearing \\
Traffic intensity in nearest road &  & Low Emission Zone (ZBE)\\
Traffic intensity in nearest major road &  & Street width (BiSCAPE)\\
                                         &  & Street canyon (BiSCAPE)\\
                                         &  & Sampler height (BiSCAPE)\\
\bottomrule
\label{Table S1}
\end{tabular}
\end{table}

\vspace{-2em}

%%%% ------------------------------------------------------%%%%
%%%% --- Table S2. Data sources for predictor variables --- %%%
%%%% ------------------------------------------------------%%%%
\begin{table}[ht]
\centering
\caption{Predictor varaibles categories created for LUR model development.}
\begin{tabular}{lll}
\toprule
\textbf{Predictor variable} & \textbf{Year} & \textbf{Source} \\
\midrule
Road network and traffic data  & 2014 & Departament de mobilitat, Ajuntament de Barcelona \\
Land uses & 2018 & Urban atlas, Copernicus \\
Population & 2016 & Institut d'Estadística de Catalunya \\
Buildings height & 2009 & Modelo Digital de Superficies Edificación  - MDSn2.5 (MDSNE)\\
Imperviousness & 2018 & Copernicus \\
Slope & 2012 & Institut Cartográfic de Catalunya \\
ZBE & 2012 & OSM\\
\bottomrule
\label{Table S2}
\end{tabular}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Deseasonalizing air pollution data --- %%% 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\subsection{Deseasonalizing pollution data}

For the development of LUR models data collected at different points of time need to be deseasonalized (process explained below) before being averaged to represent the average value of the baseline period.

\subsubsection{NO\textsubscript{2} deseasonalization}

For the NO\textsubscript{2} data (BiSCAPE + BiSC) the \textit{difference method} was used for deseasonalisation.

\begin{enumerate}
    \item Define the baseline period, that is, the total period in which the air pollution measurements were done. For NO\textsubscript{2} the baseline period was from 23\textsuperscript{rd} October 2018 to 16\textsuperscript{th} February 2022.
    \item Collect hourly air pollution data from a routine monitoring site covering the baseline period. We used data from Palau Reial monitoring station, as this is the reference site for the BiSCAPE campaigns and the BiSC cohort. NO\textsubscript{2}, data at Palau Reial was missing for only 5 \% of the period. Even though the missing rate is low, we imputed missing data by doing a linear regression with NO\textsubscript{2} levels at the monitoring site of Sants (r = 0.84). After recovering those data, the percentage of missing data at Palau Reial decreased to 0.68 \%. 
    \item Calculate the average concentration of NO\textsubscript{2} for the routine monitoring site covering the baseline measurement period: C\textsubscript{PR}(avg).
    \item Calculate the concentrations of the pollutant at the routine monitoring site for the exact same period (t) as the sample collected at the BiSC and BiSCAPE locations in the different periods: CP\textsubscript{PR}(t). 
    \item Calculate for the routine monitoring site for each sample the difference between the concentration (C\textsubscript{PR}) and the average covering the baseline measurement period: DC\textsubscript{PR}(t) = C\textsubscript{PR}(t) - C\textsubscript{PR}(avg).
    \item Calculate for each sample at site i (C\textsubscript{i}(t)) the adjusted concentration (C\textsubscript{i, adj}(t)) by computing: C\textsubscript{i,adj}(t) = C\textsubscript{i}/DC\textsubscript{PR}(t).
\end{enumerate}

\subsubsection{PM\textsubscript{2.5}, PM\textsubscript{2.5} components (Fe, Cu, Zn), and BC deseasonalization}

For the PM\textsubscript{2.5} data (BiSCAPE + BiSC) the \textit{difference method} was used for deseasonalisation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Performance metrics --- %%% 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\subsection{Performance metrics}
The following metrics were used to evaluate the LUR model for all the air pollutants assessed.

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Model development --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Model development}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Model performance --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Model performance}
The performance of the final models obtained for NO\textsubscript{2}, PM\textsubscript{2.5}, BC and PM\textsubscript{2.5 - Fe}, PM\textsubscript{2.5 - Cu}, PM\textsubscript{2.5 - Zn} is described in \textbf{\Cref{Table S3}}.

%%% --- Table with performance metrics for Hybrid models --- %%%
\begin{table}[ht]
\centering
\begin{threeparttable}
\caption{Performance metric for LUR models.}
\label{Table S3}
\begin{tabular}{llllll}
\toprule
\textbf{Pollutant} & \textbf{Year} & \textbf{N}  &  \textbf{R\textsuperscript{2}\textsubscript{10-CV}} & \textbf{RMSE\textsubscript{10-CV}} & \textbf{Predictor variables}\tnote{1} \\
\midrule
\multirow{2}{*}{NO\textsubscript{2}} & \multirow{2}{*}{2018-2021} & \multirow{2}{*}{489} & \multirow{2}{*}{0.62} & \multirow{2}{*}{4.03} & trafload25 + sqralt + majorroadlength50 + \\ 
 & & & & & roadlength25 + majorroadlength300 \\
\midrule
PM\textsubscript{2.5} & 2019 & 34 & 0.45 & 1.47 & hdres500 + trafnear + ZBE \\
\midrule
BC & 2019 & 30  & 0.83 & 0.18 & hdres50 + linesnear + pop300 + trafload + roads500 \\
\midrule
PM\textsubscript{2.5 - Fe} & 2019 & 34 & 0.89 & 0.03 & trafnear + lat + pop + ZBE  \\
\midrule
PM\textsubscript{2.5 - Cu} & 2019 & 31 & 0.87 & 0.72 & trafnear + roads1000 + ind1000 + pop25 \\
\midrule
PM\textsubscript{2.5 - Zn} & 2019 & 31  & 0.85 & 6.99 & ZBE + roads50 + ind1000 + build25  \\
\bottomrule
\end{tabular}
\begin{tablenotes}
\small
\item[1] Predictor variables (roadlength: total road length (m), majorroadlength: total major road length (m), trafload: total traffic intensity (veh/day), trafnear: traffic intensity at the nearest road (veh/day), linesnear: number traffic lines on nearest street, ZBE: Low Emissions Zone (Yes/No, ref value=No), hdres: high-density residential area (m2), roads: roads surface area (m2), ind: industry area (m2), pop: population density (inhabitants), build: building area (m2), lat: latitude (m), sqralt: squared root altitude (m)).
\end{tablenotes}
\end{threeparttable}
\end{table}

The best models were selected based on its better validation performance and on the availability of the sampling campaign that could better represent the study population.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Back-extrapolation procedure --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\subsection{Back-extrapolation procedure}

In the BiSC cohort study, we aim to study the effect of air pollution exposure using different windows of exposure (i.e., entire pregnancy, trimesters, and weeks of pregnancy). For each subject, back-extrapolated concentrations for different time periods can be calculated using daily concentrations from the background monitoring station. The station used was Palau Reial, the spatial location and the daily temporal concentration variability are shown for \textit{NO\textsubscript{2}}, \textit{PM\textsubscript{2.5}}, \textit{BC} in \textbf{\Cref{FigureS1}}.

%%% --- Figure. Monitoring stations and time-series used for LUR Back-Exptrapolation --- %%%

% Redefine the figure numbering to include 'S' for Supplementary
\renewcommand{\thefigure}{S\arabic{figure}}

%%% Figure: Monitoring stations and time-series used for LUR Back-Extrapolation %%%
\captionsetup[figure]{skip=6pt} % Adjusting the space between the figure and its caption
\begin{figure}[!htb]
    \centering
    \includegraphics[width=1.0\textwidth]{figures/location_ts_stations.png} % Adjust the path to your image
    \caption{Spatial location and daily temporal concentrations at Palau Reial station.}
    \label{FigureS1} % Label for referencing
\end{figure}

@@

\subsection{Performance metrics}
The following metrics were used to evaluate the HM for all the air pollutants assessed.

@@

%%% --- Table with performance metrics for Hybrid models --- %%%
\begin{table}[ht]
\caption{Performance metric for hybrid models.}
\centering
\begin{threeparttable}
\begin{tabular}{llllll}
\toprule
\textbf{Pollutant} & \textbf{Year} & \textbf{N} & \textbf{R\textsuperscript{2}\textsubscript{10-CV}} & \textbf{RMSE\textsubscript{10-CV}} & \textbf{Predictors}\tnote{2} \\
\midrule
\multirow{6}{*}{NO\textsubscript{2}} & \multirow{6}{*}{2018-2021} & \multirow{6}{*}{1232} & \multirow{6}{*}{0.64} & \multirow{6}{*}{7.48} & no2\_dispersion +  idw\_no2\_monitoring\_station + \\
 & & & & & distinvmajor1 + majorroadlength100 + \\
 & & & & & majorroadlength50 + trafload25 + trafnear + \\ 
 & & & & & avg\_traffic\_stations + avg\_solar\_radiation +\\ 
 & & & & & sqralt + ldres1000 +ldres500 + hdres1000 +  \\
 & & & & & build1000 + roads100 + lat + lon \\
\midrule
\multirow{4}{*}{PM\textsubscript{2.5}} & \multirow{4}{*}{2021} & \multirow{4}{*}{161} & \multirow{4}{*}{0.66} & \multirow{4}{*}{3.45} & pm25\_dispersion + idw\_pm25monitoring\_stations + \\
& & & & & avg\_traffic\_stations + ldres1000 + build25 +  pop1000 + \\
& & & & & roads25 + sqralt +  majorroadlength25 + \\
& & & & & trafload300 + build\_height\_25 + lat + lon \\
\midrule
BC & 2021 & 74 & 0.86 &  0.23 & bc\_Dispersion + idw\_nox\_monitoring\_stations +
 avg\_atmospheric\_pressure + avg\_bc\_palau\_reial +  pop100 + hdres300 + roads25 + avg\_wind\_speed + roadlength25 +  hdres50 + distinvmajor1 + avg\_traffic\_stations +  linesnear +  pop300 + 
 trafload500 + roads500
 \\
\midrule
PM\textsubscript{2.5 - Fe} & 2021 & 161 & 0.54 & 0.08 & pm25\_road\_nonexhaust +idw\_pm25monitoring\_stations + roads300 + roads1000 + ldres500 + sqralt + zbe + lat + lon \\
\midrule
PM\textsubscript{2.5 - Cu} & 2021 & 154 & 0.70 & 2.15 & pm25\_road\_nonexhaust + idw\_pm25monitoring\_stations +  linesnear + industry1000 + roads300 + ldres1000 + roads1000 + avg\_traffic\_stations + zbe + lat + lon \\
\midrule
PM\textsubscript{2.5 - Zn} & 2021 & 141 & 0.44 & 27.7 & pm25\_road\_nonexhaust + pm25\_background + idw\_pm25monitoring\_stations + linesnear + industry1000 + roads50 + build25 + roads300 + ldres1000 + roads1000 + avg\_traffic\_stations +  trafmajor +trafmajorload1000 +trafload1000 + hdres25 + zbe + lat + lon \\
\bottomrule
\end{tabular}
\begin{tablenotes}
\small
\item [2] Predictor variables (no2\_Dispersion: dispersion estimates (ug/m$^3$), pm25\_Dispersion: dispersion estimates PM2.5 (ug/m$^3$), BC\_Dispersion: dispersion estimates BC (ug/m$^3$), PM25\_road\_nonexhaust: dispersion estimates for nonexhaust PM2.5 (ug/m$^3$), PM25\_background: dispersion estimates for background PM2.5 (ug/m$^3$), idw\_no2\_monitoring\_station: weekly NO$_2$ inverse distance weighting interpolation estimates from XVCPA (ug/m$^3$), idw\_pm25\_monitoring\_station: weekly PM2.5 inverse distance weighting interpolation estimates from XVCPA (ug/m$^3$), idw\_nox\_monitoring\_station: weekly NOx inverse distance weighting interpolation estimates from XVCPA (ug/m$^3$), avg\_bc\_palau\_reial: weekly BC average concentration from Palau Reial monitoring station (ug/m$^3$), avg\_traffic\_stations: weekly average count of vehicles in AMB (count), avg\_solar\_radiation: weekly average solar radiation from Raval station (MJ/m$^2$), avg\_wind\_speed: weekly average wind speed from Raval station (km/h), avg\_atmospheric\_pressure: atmospheric pressure from Raval station (hPA), roadlength: total road length (m), majorroadlength: total major road length (m), trafload: total traffic intensity (veh/day), trafnear: traffic intensity at the nearest road (veh/day), linesnear: number traffic lines on nearest street, zbe: Low Emissions Zone (Yes/No, ref value=No), hdres: high-density residential area (m$^2$), roads: roads surface area (m$^2$), ind: industry area (m$^2$), pop: population density (inhabitants), build: building area (m$^2$), ldress: low density residential area (m$^2$), industry: industrial area (m$^2$), trafmajor: total major roads traffic intensity (veh/day), build\_height: averaged building height (meters), lat: latitude (m), sqralt: squared root altitude (m$^{1/2}$)).
\end{tablenotes}
\label{tablesX}
\end{threeparttable}
\end{table}

\newpage
\subsection{Variable importance Random Forest}

%%% --- Figure. Comparison between measurements and predictions values in the test dataset --- %%%
% Put the figure text closer to the Figure 3
\captionsetup[figure]{skip=6pt}
% We add the figure with the study domain 
\begin{figure}[!htb]
\includegraphics[width=1.0\textwidth]{figures/fig_HM_allmodels_importance_v2.png}
\caption{Variable importance for HM for \textit{NO\textsubscript{2}}, \textit{PM\textsubscript{2.5}}, \textit{BC}, \textit{PM\textsubscript{2.5 - Fe}}, \textit{PM\textsubscript{2.5 - Cu}} and \textit{PM\textsubscript{2.5 - Zn}.}}
\label{fig2}
\end{figure}

@@

%%% --- Figure. Comparison between measurements and predictions values in the test dataset --- %%%
% Put the figure text closer to the Figure 3
\captionsetup[figure]{skip=6pt}
% We add the figure with the study domain 
\begin{figure}[!htb]
\includegraphics[width=1.0\textwidth]{figures/fig_HM_test_all_models_v2.png}
\caption{Comparison between measurements and prediction values for the hybrid models for \textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}, \textit{PM\textsubscript{2.5 - Fe}}, \textit{PM\textsubscript{2.5 - Cu}}, and \textit{PM\textsubscript{2.5 - Zn}}.}
\label{fig2}
\end{figure}
